---
title: "Effects analysis"
author: '1806082'
date: "11/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setting up packages, warning = FALSE, message = FALSE}
library(devtools)
library(zoo)
source_url("https://raw.github.com/timleeman/ST344GroupProject/main/setup.R")
```

```{r getting data, , warning = FALSE, message = FALSE}
source_url("https://raw.github.com/timleeman/ST344GroupProject/main/getdata.R")
```

```{r setting up functions}
firstdiff <- function(x) {                 
  shifted <- c(0,x[1:(length(x)-1)])
  result = x-shifted
  which_negative = which(result<0)
  result[which_negative] = NA
  return(result)
}
```


```{r pivot table on the counts of lockdown, , warning = FALSE, message = FALSE}
auravisionData %>% 
  group_by(Level)%>%
  summarise(count = n())
```

```{r setting up dataframe, , warning = FALSE, message = FALSE}
covidData1<- covidData %>% filter(country %in% chosenCountries)

covidData1<- covidData1%>% 
  mutate(national_lock = ifelse(country %in% filter(auravisionData, Level == 'National')$Country, 1, 0))%>% 
  mutate(city_lock = ifelse(country %in% filter(auravisionData, Level == 'City')$Country, 1, 0))%>%
  mutate(region_lock = ifelse(country %in% filter(auravisionData, Level %in% c('Prefecture','Province','State','Region','Regional'))$Country, 1, 0))%>%
  mutate(confirmed_per_capita = confirmed/population) %>%
  mutate(daily_confirmed = firstdiff(confirmed)) %>%
  mutate(daily_deaths = firstdiff(deaths)) %>%
  mutate(deaths_per_capita = deaths/population)%>% 
  mutate(income = factor(income, levels = c("High income","Upper middle income", "Lower middle income", "Low income")))

```

```{r effects of national lockdown on retail and recreation, , warning = FALSE, message = FALSE}
plot_1<- filter(covidData1, national_lock %in% 1) %>%
  ggplot(aes(x = date, y = gcmr_retail_recreation)) +
  geom_line(aes(y = rollmean(gcmr_retail_recreation, 7, na.pad = TRUE)), colour = 'red', alpha= 0.3)+
  geom_smooth(aes(colour = 'Coutries with National Lockdowns'), data = filter(covidData1, national_lock %in% 1), se = FALSE)+
  geom_line(aes(y = rollmean(gcmr_retail_recreation, 7, na.pad = TRUE)), colour = 'blue', alpha= 0.3, data = filter(covidData1, national_lock %in% 0))+
  geom_smooth(aes(colour = 'Countries without National Lockdowns'), data = filter(covidData1, national_lock %in% 0), se = FALSE)+
  scale_colour_manual(name = 'Legend', values = c('Coutries with National Lockdowns'='red','Countries without National Lockdowns' = 'blue' ))+
  labs(title = 'Effects of a National Lockdown on Retail and Recreation', x= "Date", y = 'Activity in Retail and Recreation Centres') + 
  scale_x_date(breaks = 'months', date_labels = '%b')+
  theme(title = element_text(size = 20, hjust = 0.5, face = 'bold'), axis.title.x = element_text(size = 20), legend.text = element_text(size = 20))

```

```{r effects of national lockdown on grocery and pharmacy, , warning = FALSE, message = FALSE}
plot_2<- filter(covidData1, national_lock %in% 1) %>%
  ggplot(aes(x = date, y = gcmr_grocery_pharmacy)) +
  geom_line(aes(y = rollmean(gcmr_grocery_pharmacy, 7, na.pad = TRUE)), colour = 'red', alpha= 0.3)+
  geom_smooth(aes(colour = 'Coutries with National Lockdowns'), data = filter(covidData1, national_lock %in% 1), se = FALSE)+
  geom_line(aes(y = rollmean(gcmr_grocery_pharmacy, 7, na.pad = TRUE)), colour = 'blue', alpha= 0.3, data = filter(covidData1, national_lock %in% 0))+
  geom_smooth(aes(colour = 'Countries without National Lockdowns'), data = filter(covidData1, national_lock %in% 0), se = FALSE)+
  scale_colour_manual(name = 'Legend', values = c('Coutries with National Lockdowns'='red','Countries without National Lockdowns' = 'blue' ))+
  labs(title = 'Effects of a National Lockdown on Groceries and Pharmacies', x= "Date", y = 'Activity in Groceries and Pharmacies')+ 
  scale_x_date(breaks = 'months', date_labels = '%b')+
  theme(title = element_text(size = 20, hjust = 0.5, face = 'bold'), axis.title.x = element_text(size = 20), legend.text = element_text(size = 20))

```

```{r arranging the plots, fig.height = 10, fig.width= 15, warning = FALSE, message = FALSE}
plot_final<- ggarrange(plot_1, plot_2,ncol = 2, nrow = 1)

plot_final
```
From the 2 plots above, we can see that with the presence of a national lockdown on our list of chosen countries, the level of activity in gcmr_retail_recreation and gcmr_grocery_pharmacy have decreased dramatically compared to countries without a lockdown. This suggests that a national lockdown does hamper economic activity and thus affects the businesses by decreasing the number of customers.
                         
```{r glm for gcmr_reatail_recreation}
minModel <- glm(data = na.omit(covidData1), gcmr_retail_recreation ~ 1)
maxModel <- glm(data = na.omit(covidData1), gcmr_retail_recreation ~national_lock + confirmed_per_capita + daily_deaths + deaths_per_capita + daily_confirmed)

autoBack <- step(maxModel, direction = "backward", scope = list("lower" = minModel))
autoForward <- step(minModel, direction = "forward", scope = list("upper" = maxModel))
autoBoth <- step(minModel, direction = "both", scope = list("lower" = minModel, "upper" = maxModel))

summary(autoBack)
summary(autoForward)
summary(autoBoth)
```

```{r glm for gcmr_grocery_pharmacy}
minModel <- glm(data = na.omit(covidData1), gcmr_grocery_pharmacy ~ 1)
maxModel <- glm(data = na.omit(covidData1), gcmr_grocery_pharmacy ~national_lock + confirmed_per_capita + daily_deaths + deaths_per_capita + daily_confirmed)

autoBack <- step(maxModel, direction = "backward", scope = list("lower" = minModel))
autoForward <- step(minModel, direction = "forward", scope = list("upper" = maxModel))
autoBoth <- step(minModel, direction = "both", scope = list("lower" = minModel, "upper" = maxModel))

summary(autoBack)
summary(autoForward)
summary(autoBoth)
```


